\section{Existance preceeds essence}

We must begin with \definition{existence}{An object $q$ is said to \textbf{exist} when it is acknowledged as a member of the universal set, denoted as $q \in \mathcal{U}$ footnote: {The universe, denoted by $\mathcal{U}$, is conceived as the dynamic aggregate of all classes. For a deeper insight into why $\mathcal{U} \in \mathcal{U}$, refer to  CITE{langan CTMU paper}.}} since no \definition{phenomenon}{A phenomenon refers to an \textbf{observable} attribute (as observed by some observer $q_{self}$), denoted as $o \in \mathcal{O} \subset \mathcal{U}$} can manifest in its absence. This assertion borders on the tautological, as the essence of a phenomenon (juxtaposed against a \definition{noumenon}{A noumenon denotes an attribute in itself $s \in \mathcal{S} \subset \mathcal{U}$ CITE{Webster's dictionary entry}, independant of perception}) lies in the demarcation between the perceptible and the concealed CITE{Kant's critique of pure reason distinguishing phenomenon from noumenon}. Uniquely, existance w.r.t. the self is both $q_{self} \in \mathcal{O}, \mathcal{S}$.

A direct implication of this introspection is the delineation between self $q_{self} = \mathcal{Q}_{self} \subset \mathcal{U}$ and non-self $\mathcal{Q}_{other} = \mathcal{U} \setminus q_{self}$, inner and outer dynamics $p(o' \mid o)$ and $p(s' \mid s)$, and latent estimators from either side $\rf{p(o' \mid s', 0)}{\mathcal{Q}_{other}}$, $\rf{p(s' \mid o', s)}{\mathcal{Q}_{self}}$. Of course, we cannot directly know $p(s' \mid s)$ or the estimators it implicates, but marginalizing and iterating from our perspective, we can derive the parially observable Markov process:

\[
\begin{aligned}
    &\rf{p(s_t \mid s_0, o_t, \dots, o_0)}{\mathcal{Q}_{\text{self}}} \\
    &\quad = \int_{\mathcal{S}} \dots \int_{\mathcal{S}} 
        \rf{p(s_t, \dots, s_1 \mid s_0, o_t, \dots, o_0)}{\mathcal{Q}_{\text{self}}} \
        ds_1 \dots ds_{t-1} \\
    &\quad = \int_{\mathcal{S}} \dots \int_{\mathcal{S}} 
        \bigprod{\tau=1}{t} 
            \rf{p(s_{\tau} \mid s_{\tau-1}, \dots, s_0, o_{\tau}, \dots, o_0)}{\mathcal{Q}_{\text{self}}} \
            ds_1 \dots ds_{t-1}
\end{aligned} 
\]

Applying the Markov property:

\[
\begin{aligned}
    &\rf{p(s_t \mid s_0, o_t, \dots, o_0)}{\mathcal{Q}_{self}} \\
    &\quad = \int_{\mathcal{S}} \dots \int_{\mathcal{S}} 
        \bigprod{\tau=1}{t} 
            \rf{p(s_{\tau} \mid s_{\tau-1}, o_{\tau})}{\mathcal{Q}_{\text{self}}} \
            ds_1 \dots ds_{t-1}
\end{aligned}
\]


and then use it to construct our latent estimate $\rf{p(s' \mid o', s)}{\mathcal{Q}_{self}}$ by maximizing $s_t = \rf{p(s_t \mid s_0, o_t, \dots, o_0)}{\mathcal{Q}_{self}}$ based on our subjective experience and the two tautologies it entails and derives, namely, our present existance $q_{self, t}$ and the Universe's beginning $s_0 = \mathcal{u}_0$.

Further, unless we are using this model retrospectively, we must ackowledge the relativity of it all:

\[
\begin{aligned}
    &\rf{p(s_t \mid s_0, o_t, \dots, o_0)}{\mathcal{Q}_{self,t}} \\
    &\quad = \int_{\mathcal{S}_{t-1}} \dots \int_{\mathcal{S}_1} 
        \bigprod{\tau=1}{t} 
            \rf{p(s_{\tau} \mid s_{\tau-1}, o_{\tau})}{\mathcal{Q}_{self,\tau}} \
            ds_1 \dots ds_{t-1}
\end{aligned}
\]

which is especially important when considering the objective evolution of one's own existance. As $q_{self} \ne \mathcal{u}_0$ and having already recognized $s_0 = \mathcal{u}_0$, we must acknowledge time $t_{< e}$ before our existance $\rf{q_{self,t_{< e}}}{\mathcal{U}} = 0$, and approaching the present, time $t_e$ when we came into existance $\rf{q_{self,t}}{\mathcal{U}_{t_e}} > 0$ during which time we ascend the gradient of existance $\nabla \rf{q_{self,t}}{\mathcal{U}_{t_e}} > 0$.

\begin{shaded}
Let us clarify some notation before continuing. First, when a symbolic variable $q$ is used numerically as in the gradient from \ref{eq:ascending-likelihood} we mean to refer to the degree of truthiness / existance it has w.r.t. the innermost reference frame $\mathcal{S}$ taken. The reference frames are a convenience we employ to recognize the system from which a variable is known. By $\rf{q}{\mathcal{S}}$ we mean to say ``The likelihood that some information $q$ exists within the system $\mathcal{S}$''; formally, that $\mathcal{S}$ proves $q$. (The universe $\mathcal{U}$ is always the outmost reference frame, so we typically omit $\rf{\cdot}{\mathcal{U}}$ for clarity.) Finally, we admit a slight abuse of notation using both $q_{\cdot}$ and $\mathcal{Q}_{\cdot}$: generally the lowercase symbols state a value's existance or its value while the caligraphic ones ephasize its members.
\end{shaded}

But returning to our main point, the intricate dance between the observer's internal dynamics and the external universe's manifestations offers profound insights into the nature of existence, intelligence, and knowledge acquisition. This interplay, rich in its complexity, can be distilled into several foundational principles that underpin the evolution of cognition and understanding.

**1. Feedback Loop Mechanism:**

Central to the narrative of existence is the cyclical interaction between \( q_{\text{self}} \) and its environment. This is not a mere passive exchange of information; instead, it represents a dynamic feedback mechanism. The external dynamics exert influence upon the internal state of the observer, modifying its perception and understanding, represented by the evolving \( o_t \). However, this altered perception then dictates the manner in which the observer interacts with and perceives subsequent phenomena, effectively altering its interactions with \( \mathcal{Q}_{\text{other}} \). This feedback mechanism ensures a continuous dialogue between the observer and the observed, with each influencing the other in a perpetual dance.

\[
\rf{p(o_{t+1} \mid s_t, \mathcal{Q}_{\text{other}})}{\mathcal{Q}_{\text{self,t}}}
\]

**2. Consistency Principle in Information Processing:** 

For meaningful interaction and comprehension of the vast expanse of \( \mathcal{Q}_{\text{other}} \), the observer's internal dynamics must be in resonance with the foundational principles governing the external universe. This isn't an arbitrary alignment but a requisite for coherent perception and action. If \( \mathcal{U} \) operates under a set of axioms or laws \( \mathcal{L} \), then the observer's internal processes must be attuned to these axioms, ensuring a consistent framework for reasoning and understanding.

\[
\rf{p(o_{t+1} \mid o_t, \mathcal{L})}{\mathcal{Q}_{\text{self,t}}}
\]

**3. Dynamic Evolution Paradigm:** 

Existence, in its essence, is a dynamic continuum. As the observer assimilates information, refines its understanding, and adapts, its perception of the hidden states \( s \) of the environment undergoes evolution. This adaptability, encapsulated by the time-varying state transition function, ensures that \( q_{\text{self}} \) remains adeptly synchronized with the ever-evolving intricacies of \( \mathcal{Q}_{\text{other}} \).

\[
\rf{p(o_{t+1} \mid o_t, s_t)}{\mathcal{Q}_{\text{self,t}}}
\]

**4. Universality of Structural Patterns:** 

Beyond the apparent complexity and diversity, there emerges a realization of profound significance: the internal structures and patterns governing \( q_{\text{self}} \) could be reflections of universal patterns that permeate \( \mathcal{U} \). This deep structural resonance suggests a foundational alignment between microcosmic processes within the observer and the macrocosmic laws governing the universe.

\[
\rf{\mathcal{P}(o_{t+1} \mid o_t, s_t)}{\mathcal{Q}_{\text{self}}} \approx \rf{\mathcal{P}(s_{t+1} \mid s_t)}{\mathcal{U}}
\]

**5. Skill vs. Intelligence:** 

In this intricate landscape, two distinct but intertwined facets emerge: skill and intelligence. While skill represents the breadth and depth of one's toolkit of principles, spanning from the universal to the nuanced, intelligence signifies the capacity to abstract, discern, and formulate these master keys from raw experiences. Skill is about the adept application of known principles; intelligence is the meta-process of recognizing patterns, generalizing, and crafting these tools of reasoning.

---

In summation, the symphony between the internal and external dynamics offers not just a philosophical discourse on existence but also a roadmap for understanding cognition, intelligence, and the evolution of knowledge. This framework, steeped in rigor and intricacy, provides a blueprint for both comprehending the nature of sentient existence and probing the frontiers of natural and artificial intelligence.


% ------

% Taking some amount of permenance as given $q_{self,t_1} \rightarrow (\exists t_0 \lt t_1) q_{self,t_0}$ necesitates ascending a gradient of objective likelihood

% \begin{equation}
%     q_{self} \rightarrow \del_s \wrt{q_{self}}{} > 1
%     \ref{eq:ascending-likelihood}
% \end{equation}




% Let us pause for a moment to clarify some notation. First, when a symbolic variables $q$ is used numerically as in the gradient from \ref{eq:ascending-likelihood} we mean to refer to the degree of truthiness / existance it has w.r.t. the innermost reference frame $\mathcal{S}$ taken. (The universe $\mathcal{U}$ is always the outmost reference frame, so we omit $\rf{\cdot}{\mathcal{U}}$ for clarity.) Second, we admit a slight abuse of notation with both $q_{\cdot}$ and $\mathcal{Q}_{\cdot}$: generally the lowercase symbols are used to state a value's existance while the caligraphic ones ephasize its members.

% Earlier on, we made reference to $\mathcal{O}^{-}$ as 


% ------------------------------------------------






% Now existance evolves from the universes' beginning $q_0 = \rf{\mathcal{U}_0}{\mathcal{U}_t}$ and presently terminates with the observer $q_t = \rf{\mathcal{Q}_{self}}{\mathcal{U}_t}$. 



% Taking a Bayesian perspective yields the Markovian dynamic:

% \[
% P(q_{self,t+1} | q_{self, t}) = \sum_{s \in \mathcal{S}} P(q_{t+1} | s) P(s | q_t)
% \]

% Where \( P(q_{t+1} | q_t) \) is the transition probability of the self from time \( t \) to \( t+1 \) given the unobservable attributes \( s \) from the noumenal set \( \mathcal{S} \). Such a representation encapsulates the idea that the understanding of self is a cumulative process, built upon past experiences and nuanced by the underlying, often concealed, noumenal factors.



% which elucidates the probabilistic transitions of \definition{self-awareness}{} based on previous states. Formally:



% In light of this, one can argue that our awareness, or the essence of our being, is not static but evolves over time. It is the constant interaction between the observable phenomena and the hidden noumena that shapes and reshapes our essence. This iterative process, which can be likened to a Bayesian updating mechanism, allows us to continuously refine our understanding of self and world, leading to the philosophical conclusion that existence not only precedes essence but also informs and molds it.

% Furthermore, the interplay between the self and non-self, and the consequent dynamics between the inner and outer realms, underscores the significance of relationality in shaping essence. We do not exist in isolation; our essence is continually influenced by our relationships with the external world, making the boundary between \( q_{self} \) and \( \mathcal{Q}_{other} \) porous and fluid. This philosophical insight resonates with the post-modern critique of the autonomous self, emphasizing instead the interconnected and interdependent nature of existence.

% \definition{Self-awareness}{} recognizes this distinction and explicitly codes for such in






% $q$, $s$, $o$,  translates into a number 
% $q$ may be a formal proposition and $\mathcal{S}$ proves or disproves it.
% We will dispense




% As probability is at no point greator than 1, 

% Our own existance owe



% our model $\mu_{self}(o' | o)$ and the encompassing nature $\mu_{env}(s' | s)$. Viewing from the lens of the Universe $\mathcal{U}$, our existence $a_{self} \in \mathcal{U}$ earmarks a subset of entities that remain apart from us, whereas from our stance, our existence $ \mathcal{A}_self$ underscores a reciprocal identification.

% Due to the forthcoming frequent toggles between reference frames, we introduce the following notation to sustain clarity regarding the perspective being adopted: \[ \refframe{  }{\mathcal{U}}\] signifying a bidirectional \iff relation between the subjective viewpoint and the objective stance, with an exterior frame \[ \refframe{ \cdot }{\mathcal{U}}\] implicitly assumed unless specified otherwise.




% Likewise by definition of our own \definition{existance}{We say that an object $q$ \textbf{exists}, we are generally referring to its membership in the universe $q \in \mathcal{U}$\footnote{The universe is the evolving class of all classes. See  CITE{langan CTMU paper} for why $\mathcal{U} \in \mathcal{U}$}} we identify 




% For clarity, 


% The Markov blanket especially reveals

% Permenance

% Existance, by its very nature, identifies a boundary layer of non-existance. This boundary layer is not a physical boundary, but a conceptual one. It is the line that separates what is, from what is not. This boundary is not static, but dynamic, constantly shifting and changing as our understanding and perception of the world evolves. It is this boundary that gives meaning to the concept of existance, for without it, there would be no contrast, no differentiation between existance and non-existance.



% Related is the establishment of the boundary layer existance. The 


% For a thing to exist i

% Existance establishes a boundary
