By acquiring rich formalizations, intelligence grows in complexity, as evident in the recent advances in AI.
    - Everything, everywhere, all at once. The spirit of AI programming itself into every substrate concievable. The shoggoth's tentacles penetrating every aspect of our lives.
    - but we can't formalize everything, so we must recognize the limits of formalization
        - math can only take us to places its infrastructure has already been built
            - Even as Socrates himself stated: quote about logic being for debate, not discovery
        - so why not build complextely new systems of intuition, even if they aren't perfect?
            x  I call them informalisms

Why informalize?

There's so much to explore and way too little time to explore it. [1] estimates we only have 7 ? months left until AGI is achieved.

Humans possess a diverse array of mechanisms that equip them to engineer complex patches onto open-ended problems. At the neuron level, these include a cell membrane, thousands of synapses, an axon, and an internal signal interaction network to guide the cycle of stimulus-response-stimulus. Neurons organize into highly modular heterogeneous networks, and their emergent rhythms of structural and electrochemical activity provide a slate against which external disruptions are compared to build meaningful representations. The human body provides a rich set of sensory-motor modalities, and a natural curriculum of physical, mental, and social development exposes increasingly many degrees of freedom to interaction accompanied by extrinsic and intrinsic feedback mechanisms. With sufficient social and technological extensions, humans are even able to modify their own environment, physiology, and brain.









I could give you many more examples of how to *informalize* soft concepts into hard ones, but I think you get the point. The key is to realize that we can't formalize everything, but we can formalize *something*, and that something at least gives us (or the AI) a starting point to evolve from.